R-LIME: LIME法における矩形制約と最適化

### 2024-03-02 paper/paper.tex
Work on paper

\ifnum\mode=0
	\documentclass[11pt]{article}
	\usepackage{prmlthesis}
	\usepackage[a4paper]{geometry}
\else
	\documentclass[runningheads]{llncs}
\fi

@@

\ifnum\mode=0
	\title{%
		R-LIME: Rectangular Constraints and Optimization for Local Interpretable
		Model-agnostic Explanation Methods
	}
	\etitle{}
	\author{Genji Ohara}
	\institute{hogehoge}
\else
	\title{%
		R-LIME: Rectangular Constraints and Optimization for Local Interpretable
		Model-agnostic Explanation Methods}
	% Abbreviated title for the running head
	\titlerunning{%
		R-LIME: Rectangular Constraints and Optimization for LIME Methods
	}

@@

Model-dependent methods,
such as DTD (Deep Taylor Decomposition)~\cite{montavon2017explaining}
and LRP (Layer-wise Relevance Propagation)~\cite{bach2015pixel},
primarily focus on deep neural networks and
explain the models' behavior using their parameters~\cite{samek2021explaining}.
While these methods can provide detailed explanations
(e.g., layer-wise explanations for neural networks)
using the models' internal parameters,
it is often challenging
to apply the same method to models with different structures.

On the other hand,
model-agnostic methods
do not rely on the models' parameters and utilize only their output.
Although these methods are applicable to any model,
they cannot explain the reasoning process inside the model.
Furthermore, model-agnostic methods are classified into
\emph{global} and \emph{local} methods based on their locality in input space.
Global methods,
such as PDP (Partial Dependence Plots)~\cite{friedman2001greedy}
and ALE (Accumulated Local Effects)~\cite{apley2020visualizing},
aim to explain the model's behavior across the entire input space,
seeking explanations valid for any input.
However, providing global explanations becomes challenging
as the model's complexity increases.
In contrast, local methods,
such as LIME (Local Interpretable Model-agnostic Explanations)~\cite{ribeiro2016why}
and Anchor~\cite{ribeiro2018anchors},
% SHAP~\cite{lundberg2017unified},
% EXPLAN~\cite{rasouli2020explan},
% LORE (Local Rule-based Explanations)~\cite{guidotti2018local}
% and BELLA (Black box model Explanations by Local Linear Approximations)
% \cite{radulovic2023bella},
explain output of the model for a specific input
(focal point) in the vicinity of that input.
While local methods offer more simple and accurate explanations
than global methods,
the scope of the explanation is limited to the vicinity of the focal point.

\begin{figure}[tbp]
	\def\scale{0.38}
	\centering
	\vspace{0.5cm}
	\begin{subfigure}[t]{0.55\textwidth}
		\centering
		\includegraphics[scale=\scale]{example-instance}
		\caption{Focal point.}\label{fig:example-instance}
	\end{subfigure}
	\hfill
	\begin{subfigure}[t]{0.45\textwidth}
		\begin{subfigure}[t]{\textwidth}
			\centering
			\includegraphics[scale=\scale]{example-lime}
			\caption{%
				Explanation by LIME\@.
				It provides contributions of each feature to the output,
				but does not explicitly indicate its scope.
			}\label{fig:example-lime}
			\vspace{0.4cm}
		\end{subfigure}
		\begin{subfigure}[t]{\textwidth}
			\centering
			\includegraphics[scale=\scale]{example-anchor}
			\caption{%
				Explanation by Anchor.
				It provides the effective scope of the explanation,
				but does not show the influence of each feature.
			}\label{fig:example-anchor}
		\end{subfigure}
	\end{subfigure}
	\hspace{0.3cm}
	\begin{subfigure}[t]{0.45\textwidth}
		\centering
		\vspace{-2.24cm}
		\includegraphics[scale=\scale]{example-rlime}
		\caption{%
			Explanation by R-LIME\@.
			It provides both contributions of each feature and
			its effective scope.
		}\label{fig:example-rlime}
	\end{subfigure}
	\caption[Example of explanations by LIME, Anchor and R-LIME]{%
		Example of explanations by LIME~\cite{ribeiro2016why},
		Anchor~\cite{ribeiro2018anchors} and R-LIME (our proposed method).
		for a sentiment prediction model.
	}\label{fig:example}
\end{figure}
In this paper,
we focus on local model-agnostic methods.
An example of explanations by LIME and Anchor
for a sentiment prediction model is illustrated in \cref{fig:example}.
LIME~\cite{ribeiro2016why} approximates the complex decision boundary
locally using a linear model around the given focal point,
then provides the weights of the linear model as the contribution of each feature
to the output.
The explanation by LIME (\cref{fig:example-lime}) provides suggests that
the word ``not'' mainly contributes to the positive prediction,
but does not explicitly indicate the effective scope.
Without the scope,
users might mistakenly apply the knowledge derived from the explanation
to other instances far from the focal point,
potentially leading to misunderstanding of the black-box model's behavior
\cite{ribeiro2018anchors}.
For example,
user may apply the insights derived from the explanation
to the sentence ``This book is not good.''
and mistakenly conclude that the word ``not''
mainly contributes to the positive prediction for this sentence as well,
which is incorrect obviously.
Anchor~\cite{ribeiro2018anchors} represents a rectangular region
containing the focal point,
that maximizes the probability of the black-box classifier outputting
the same label as the focal point.
While Anchor provides an effective scope of the explanation,
the information it includes is less than LIME,
limiting its utility~\cite{ribeiro2018anchors}.
The explanation by Anchor (\cref{fig:example-anchor})
suggests that changing words other than ``not'' and ``bad''
has little impact on the classifier's output.
While it clearly cannot be applied to the sentence ``This book is not good''
because of not including the word ``bad'',
the explanation does not provide details about the influence of each word,
resulting in less user insight into the model's behavior compared to LIME.

@@

\section{Conclusion}
We identified challenges in existing methods for local model-agnostic post-hoc
explanations of black-box classifiers and proposed R-LIME to address them.
We represented the rectangular region for local linear approximation as a
conjunction of feature predicates and proposed an algorithm to
maximize coverage under the constraint of minimum approximation accuracy.
Comparing the outputs of LIME and R-LIME on real-world datasets,
we demonstrated that explanations provided by R-LIME have clearer application
scopes and can be evaluated by users for reliability and generality.
However, we discussed the instability of behavior concerning imbalanced label
distributions and raised questions about the theoretical validity of using
the KL-LUCB algorithm.

>>>>

\ifnum\mode=0
	\documentclass[11pt]{article}
	\usepackage{luatexja}
	\usepackage{prmlthesis}
	\usepackage[a4paper]{geometry}
\else
	\documentclass[runningheads]{llncs}
\fi

@@

\ifnum\mode=0
	\title{%
		R-LIME: LIME法における矩形制約と最適化
	}
	\etitle{%
		R-LIME: Rectangular Constraints and Optimization for Local Interpretable
		Model-agnostic Explanation Methods
	}
	\author{大原 玄嗣 (Genji Ohara)}
	\institute{%
		北海道大学 工学部 \\
		情報エレクトロニクス学科 \\
		情報理工学コース \\
		情報認識学研究室
	}
	\date{2024年3月}
	% \institute{%
	% 	Laboratory for Pattern Recognition and Machine Learning,\\
	% 	Division of Computer Science and Information Technology,\\
	% 	Department of Electronics and Information Engineering,\\
	% 	School of Engineering, Hokkaido University
	% }
\else
	\title{%
		R-LIME: Rectangular Constraints and Optimization for Local Interpretable
		Model-agnostic Explanation Methods}
	% Abbreviated title for the running head
	\titlerunning{%
		R-LIME: Rectangular Constraints and Optimization for LIME Methods
	}

@@

Model-dependent methods,
such as Deep Taylor Decomposition (DTD)~\cite{montavon2017explaining}
and Layer-wise Relevance Propagation (LRP)~\cite{bach2015pixel},
primarily focus on deep neural networks and
explain the models' behavior using their parameters~\cite{samek2021explaining}.
While these methods can provide detailed explanations
(e.g., layer-wise explanations for neural networks)
using the models' internal parameters,
it is often challenging
to apply the same method to models with different structures.

On the other hand,
model-agnostic methods
do not rely on the models' parameters and utilize only their output.
Although these methods are applicable to any model,
they cannot explain the reasoning process inside the model.
Furthermore, model-agnostic methods are classified into
\emph{global} and \emph{local} methods based on their locality in input space.
Global methods,
such as Partial Dependence Plots (PDP)~\cite{friedman2001greedy}
and Accumulated Local Effects (ALE)~\cite{apley2020visualizing},
aim to explain the model's behavior across the entire input space,
seeking explanations valid for any input.
However, providing global explanations becomes challenging
as the model's complexity increases.
In contrast, local methods,
such as LIME (Local Interpretable Model-agnostic Explanations)~\cite{ribeiro2016why}
and Anchor~\cite{ribeiro2018anchors},
% SHAP~\cite{lundberg2017unified},
% EXPLAN~\cite{rasouli2020explan},
% LORE (Local Rule-based Explanations)~\cite{guidotti2018local}
% and BELLA (Black box model Explanations by Local Linear Approximations)
% \cite{radulovic2023bella},
explain output of the model for a specific input
(focal point) in the vicinity of that input.
While local methods offer more simple and accurate explanations
than global methods,
the scope of the explanation is limited to the vicinity of the focal point.

\begin{figure}[tbp]
	\def\scale{0.38}
	\centering
	\vspace{0.5cm}
	\begin{subfigure}[t]{0.55\textwidth}
		\centering
		\includegraphics[scale=\scale]{example-instance}
		\caption{Focal point.}\label{fig:example-instance}
	\end{subfigure}
	\hfill
	\begin{subfigure}[t]{0.45\textwidth}
		\begin{subfigure}[t]{\textwidth}
			\centering
			\includegraphics[scale=\scale]{example-lime}
			\caption{%
				Explanation by LIME\@.
				It provides contributions of each feature to the output,
				but does not explicitly indicate its scope.
			}\label{fig:example-lime}
			\vspace{0.4cm}
		\end{subfigure}
		\begin{subfigure}[t]{\textwidth}
			\centering
			\includegraphics[scale=\scale]{example-anchor}
			\caption{%
				Explanation by Anchor.
				It provides the effective scope of the explanation,
				but does not show the influence of each feature.
			}\label{fig:example-anchor}
		\end{subfigure}
	\end{subfigure}
	\hspace{0.3cm}
	\begin{subfigure}[t]{0.45\textwidth}
		\centering
		\ifnum\mode=0
			\vspace{-1.79cm}
		\else
			\vspace{-2.24cm}
		\fi
		\includegraphics[scale=\scale]{example-rlime}
		\caption{%
			Explanation by R-LIME\@.
			It provides both contributions of each feature and
			its effective scope.
		}\label{fig:example-rlime}
	\end{subfigure}
	\caption[Example of explanations by LIME, Anchor and R-LIME]{%
		Example of explanations by LIME~\cite{ribeiro2016why},
		Anchor~\cite{ribeiro2018anchors} and R-LIME (our proposed method).
		for a sentiment prediction model.
	}\label{fig:example}
\end{figure}
In this paper,
we focus on local model-agnostic methods.
An example of explanations by LIME and Anchor
for a sentiment prediction model is illustrated in \cref{fig:example}.
LIME~\cite{ribeiro2016why} approximates the complex decision boundary
locally using a linear model around the given focal point,
then provides the weights of the linear model as the contribution of each feature
to the output.
The explanation by LIME (\cref{fig:example-lime}) provides suggests that
the word ``not'' mainly contributes to the positive prediction,
but does not explicitly indicate the effective scope.
Without the scope,
users might mistakenly apply the knowledge derived from the explanation
to other instances far from the focal point,
potentially leading to misunderstanding of the black-box model's behavior
\cite{ribeiro2018anchors}.
For example,
user may apply the insights derived from the explanation
to the sentence ``This book is not good.''
and mistakenly conclude that the word ``not''
mainly contributes to the positive prediction for this sentence as well,
which is incorrect obviously.
Anchor~\cite{ribeiro2018anchors} represents a rectangular region
containing the focal point,
that maximizes the probability of the black-box classifier outputting
the same label as the focal point.
While Anchor provides an effective scope of the explanation,
the information it includes is less than LIME,
limiting its utility~\cite{ribeiro2018anchors}.
The explanation by Anchor (\cref{fig:example-anchor})
suggests that changing words other than ``not'' and ``bad''
has little impact on the classifier's output.
While it clearly cannot be applied to the sentence ``This book is not good''
because of not including the word ``bad'',
the explanation does not provide details about the influence of each word,
resulting in less user insight into the model's behavior compared to LIME.

@@

\section{Conclusion}
We identified challenges in existing methods for local model-agnostic post-hoc
explanations of black-box classifiers and proposed R-LIME to address them.
We represented the rectangular region for local linear approximation as a
conjunction of feature predicates and proposed an algorithm to
maximize coverage under the constraint of minimum approximation accuracy.
Comparing the outputs of LIME and R-LIME on the real-world dataset,
we demonstrated that R-LIME provides a clear application scope of the
explanation and can be evaluated by users for its reliability and generality.
However, we discussed the instability of behavior concerning imbalanced label
distributions and raised questions about the theoretical validity of using
the KL-LUCB algorithm.
